---
title: "Fit Cavity Pressure"
author: "Al Cooper"
date: "11/5/2020"
output: pdf_document
---

```{r initialization,echo=FALSE,include=FALSE}

library(knitr)
options(warn=-1)
opts_chunk$set(echo=FALSE, include=FALSE, fig.lp="fig:", background='#F5F2F8')
opts_chunk$set(fig.width=6, fig.height=5, fig.align="center", digits=4)
options(digits=4)
thisFileName <- "DPpressure"
library(Ranadu, quietly = TRUE, warn.conflicts=FALSE)
SaveRData <- sprintf("%s.Rdata", thisFileName)

```

\section*{Background}

The study reported in the algorithm note ``DP\_Pressure\_Correction.pdf''
applied to the GV and found a representation different from that being
used in QAtools even for the GV. It appears useful to revisit this
issue for both aircraft.

\section*{Analysis}

\subsection*{C-130}

For the C-130, all the research flights from WECAN were used except
rf14, which appeared to provide some outlier values compared to other
flights. The differences CAVP\_DPT-PSXC and CAVP\_DPB-PSXC were calculated
and an empirical search for reasonable fits led to a reasonable representation
of both in terms of QCXC. MACHX, and AKRD where QCXC was the leading
contribution to the reduction of variance for CAVP\_DPT-PSXC but AKRD
was by far the most important for CAVP\_DPB-PSXC. The residual error
for both fits was about 1 hPa. Furthermore, for the C-130 the measured
pressures differ little (typcally <10~hPa) from the ambient pressure
so the corrections for these pressure measurements are small.

```{r get-data, echo=FALSE, include=TRUE, fig.height=4}

VarList <- c("CAVP_DPT", "CAVP_DPB", "PSXC", "ATX", "DPXC", "QCXC", 
  "EWX", "ACINS","GGLAT", "TASX", "ROLL", "AKRD", "SSRD", "MACHX") 
Data <- getProjectData('WECAN', standardVariables(VarList))
Data <- Data[Data$RF != 14, ]

Data$DP1 <- Data$CAVP_DPT - Data$PSXC
Data$DP2 <- Data$CAVP_DPB - Data$PSXC
Data <- dplyr::filter(Data, TASX > 80, abs(ROLL) < 2)
f1 <- lm(DP1 ~ QCXC + MACHX + AKRD, data=Data)
f2 <- lm(DP2 ~ QCXC + MACHX + AKRD, data=Data)
cf1 <- coef(f1)
cf2 <- coef(f2)
Data$CAVP1 <- Data$PSXC + cf1[1] + cf1[2] * Data$QCXC + cf1[3] * Data$MACHX +
              cf1[4] * Data$AKRD
Data$CAVP2 <- Data$PSXC + cf2[1] + cf2[2] * Data$QCXC + cf2[3] * Data$MACHX +
              cf2[4] * Data$AKRD
Data$DELP1 <- Data$CAVP_DPT - Data$CAVP1
Data$DELP2 <- Data$CAVP_DPB - Data$CAVP2
Data %>% dplyr::select(DELP1, RF) %>% 
    binStats() %>%
    ggplot(aes(x=xc)) + 
    geom_point(aes(y=ybar), color='blue', na.rm=TRUE) + 
    geom_errorbar(aes(ymin=ybar-sigma, ymax=ybar+sigma)) +
    ylim(c(-5,5)) + ylab('CAVP_DPT-fit') + xlab('WECAN flight number') +
    theme_WAC() 

Data %>% dplyr::select(DELP2, RF) %>% 
    binStats() %>%
    ggplot(aes(x=xc)) + 
    geom_point(aes(y=ybar), color='forestgreen', na.rm=TRUE) + 
    geom_errorbar(aes(ymin=ybar-sigma, ymax=ybar+sigma)) +
    ylim(c(-5,5)) + ylab('CAVP_DPB-fit') + xlab('WECAN flight number') +
    theme_WAC() 
#print(cf1)
#print(cf2)

```

\subsection*{GV:}

The same approach was used for the GV, using the data from the ORCAS
project, selected because it was recent and yet covered a wider range
in altitude than SOCRATES. Only ORCAS flights 9--19 were included
because there seemed to be a shift before this and these appeared
more consistent. In this case, the measured pressures were significantly
different from the ambient pressures and the corrections for the measured
pressures are quite important.

```{r get-data2, echo=FALSE, include=TRUE, fig.height=4}

VarList <- c("CAVP_DPR", "CAVP_DPL", "PSXC", "ATX", "DPXC", "QCXC", 
  "EWX", "ACINS","GGALT", "TASX", "ROLL", "AKRD", "SSRD", "MACHX") 
Data <- getProjectData('ORCAS', standardVariables(VarList))
Data <- Data[Data$RF > 9, ]


Data$DP1 <- (Data$CAVP_DPL - Data$PSXC) / Data$PSXC
Data$DP2 <- (Data$CAVP_DPR - Data$PSXC) / Data$PSXC
Data <- dplyr::filter(Data, TASX > 110, abs(ROLL) < 2)
f1g <- lm(DP1 ~ QCXC + MACHX + AKRD, data=Data)
f2g <- lm(DP2 ~ QCXC + MACHX + AKRD, data=Data)
cf1g <- coef(f1g)
cf2g <- coef(f2g)
Data$CAVP1 <- Data$PSXC * (1 + cf1g[1] + cf1g[2] * Data$QCXC + 
    cf1g[3] * Data$MACHX + cf1g[4] * Data$AKRD)
Data$CAVP2 <- Data$PSXC * (1 + cf2g[1] + cf2g[2] * Data$QCXC + 
    cf2g[3] * Data$MACHX + cf2g[4] * Data$AKRD)
Data$DELP1 <- Data$CAVP_DPL - Data$CAVP1
Data$DELP2 <- Data$CAVP_DPR - Data$CAVP2
Data %>% dplyr::select(DELP1, RF) %>% 
    binStats(bins=11) %>%
    ggplot(aes(x=xc)) + 
    geom_point(aes(y=ybar), color='blue', na.rm=TRUE) + 
    geom_errorbar(aes(ymin=ybar-sigma, ymax=ybar+sigma)) +
    ylab('CAVP_DPL-fit') + xlab('ORCAS flight number') +
    theme_WAC() 

Data %>% dplyr::select(DELP2, RF) %>% 
    binStats(bins=11) %>%
    ggplot(aes(x=xc)) + 
    geom_point(aes(y=ybar), color='forestgreen', na.rm=TRUE) + 
    geom_errorbar(aes(ymin=ybar-sigma, ymax=ybar+sigma)) +
    ylab('CAVP_DPR-fit') + xlab('ORCAS flight number') +
    theme_WAC() 

```

\section*{Recommendations}

For the C-130, the cavity pressures are represented very well (within
about 1 hPa) by these equations:\\
\[
\mathrm{CAVP\_DPT=}\mathrm{PSXC}+a_{0}+a_{1}\mathrm{QCXC+a_{2}\mathrm{MACHX+a_{3}\mathrm{AKRD}}}
\]
\[
\mathrm{CAVP\_DPB=}\mathrm{PSXC}+b_{0}+b_{1}\mathrm{QCXC+b_{2}\mathrm{MACHX+b_{3}\mathrm{AKRD}}}
\]
where \{$a_{0-3}$\}=\{`r round(cf1[1], 4)`, `r format(cf1[2], digits=4, nsmall=4)`,
`r cf1[3]`, `r cf1[4]`\} and \{$b_{0-3}$\}=\{`r cf2[1]`,
`r cf2[2]`, `r cf2[3]`, `r cf2[4]`\}.

For the GV, the cavity pressures are represented by\\
\[
\mathrm{CAVP\_DPL=}\mathrm{PSXC}*(1+c_{0}+c_{1}\mathrm{QCXC+c_{2}\mathrm{MACHX+c_{3}\mathrm{AKRD)}}}
\]
\[
\mathrm{CAVP\_DPR=}\mathrm{PSXC}*(1+d_{0}+d_{1}\mathrm{QCXC+d_{2}\mathrm{MACHX+d_{3}\mathrm{AKRD)}}}
\]
where \{$c_{0-3}$\}=\{`r cf1g[1]`, `r cf1g[2]`, `r cf1g[3]`,
`r cf1g[4]`\} and \{$d_{0-3}$\}=\{`r format(cf2g[1], digits=4, nsmall=4)`,
`r cf2g[2]`, `r format(cf2g[3], digits=4, nsmall=4)`, `r cf2g[4]`\}.
In these cases, the standard errors in the resulting representations
of the cavity pressure are respectively about `r round(sd(Data$CAVP_DPL-Data$CAVP1, na.rm=TRUE),1)`
and `r round(sd(Data$CAVP_DPR-Data$CAVP2, na.rm=TRUE),1)`~hPa,
while the average differences between the cavity pressures and the
ambient pressure are respectively about `r round(mean(Data$CAVP_DPL-Data$PSXC, na.rm=TRUE),0)`
and `r round(mean(Data$CAVP_DPR-Data$PSXC, na.rm=TRUE),0)`~hPa.
This indicates that, even though the empirical representation has
greater standard deviation than would be desirable, it is still important
to use these if the direct measurement is missing because the influence
on the measured dewpoint would be significant.




